---
title: "NorthWind Database Query Questions."
author: "ITZEL SANCHEZ"
date: "2023-04-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}


```
The Northwind database represents a fictitious company sales database
"Northwind", which imports and exports specialty foods worldwide.
Contains Following Information:

• Providers (Suppliers): information on Northwind's suppliers.

• Clients (Customers): information of clients who purchase products from Northwind.

• Employees: Northwind employee information.

• Products (Products): Northwind product information.

• Carrier (Shipper): information about the courier companies used to
  send Northwind products.
  
• Orders and Order Details (Orders and Order_Detail): record of orders
  made, as well as the details of said orders.




```{r db}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), dbname = "C:/Users/anura/Desktop/Rey Castro/Northwind.sqlite")

```

1.Obtain a table with the products and units in Stock of the products that are discontinued. Sort descending by Stock and show the first 5 records.


```{r}
query <- "SELECT ProductName, UnitsInStock 
          FROM Product 
          WHERE Discontinued = 1 
          ORDER BY UnitsInStock DESC 
          LIMIT 5"
result <- dbGetQuery(con, query)

# Display the result as a table
table(result)
```

2. Build a table with the product name, unit price, and quantity of products bought. Also, add a column where the total sales are calculated.
```{r}
sales_table <- dbGetQuery(con, "
SELECT Product.ProductName, Product.UnitPrice, SUM(OrderDetail.Quantity) AS Quantity, 
       SUM(OrderDetail.Quantity * OrderDetail.UnitPrice) AS TotalSales
FROM Product
INNER JOIN OrderDetail ON Product.ID = OrderDetail.ProductID
INNER JOIN 'Order' ON OrderDetail.OrderID = 'Order'.ID
GROUP BY Product.ProductName
ORDER BY TotalSales DESC")

# Display the result as a table
sales_table
```
3. Show total sales by customer's country of origin.

```{r}
#Query the database to obtain the desired table
```


```{r}
query <- "SELECT c.Country, SUM(od.Quantity * od.UnitPrice) AS TotalSales
          FROM Customer c
          JOIN [Order] o ON c.Id = o.CustomerId
          JOIN OrderDetail od ON o.Id = od.OrderId
          GROUP BY c.Country"

# Retrieve the data and store it in a data frame
results <- dbGetQuery(con, query)



results
```


```{r}
```
4. Build a table with the product name, unit price, and quantity of products bought. Also, add a column where the total discounts are calculated.

```{r}

query <- "SELECT Product.ProductName, Product.UnitPrice, SUM(OrderDetail.Quantity) as TotalQuantity, 
       SUM(OrderDetail.Discount) as TotalDiscount, 
       SUM((Product.UnitPrice * OrderDetail.Quantity) * (1 - OrderDetail.Discount)) as TotalSales
FROM Product
INNER JOIN OrderDetail ON Product.Id = OrderDetail.ProductId
GROUP BY Product.ProductName, Product.UnitPrice"

tables <- dbGetQuery(con, query)

tables
```

5. Calculate the total number of discounts by product category.
```{r}
query <- "SELECT Category.CategoryName, SUM(OrderDetail.Discount) as TotalDiscount
FROM Category
INNER JOIN Product ON Category.ID = Product.CategoryId
INNER JOIN OrderDetail ON Product.ID = OrderDetail.ProductID
GROUP BY Category.CategoryName"

tables <- dbGetQuery(con, query)
tables
```
6. Suppliers in Spain request to know the demand for their products through the months. The information that will be shared corresponds to the amount of product sold from according to your order date. Show the information graphically.
```{r}
query <- "SELECT Product.ProductName, strftime('%m', [Order].OrderDate) as Month, SUM(OrderDetail.Quantity) as TotalQuantity
FROM [Order]
INNER JOIN OrderDetail ON [Order].ID = OrderDetail.OrderID
INNER JOIN Product ON OrderDetail.ProductID = Product.ID
INNER JOIN Supplier ON Product.SupplierID = Supplier.ID
WHERE Supplier.Country = 'Spain'
GROUP BY Product.ProductName, Month
ORDER BY Month;"

tables <- dbGetQuery(con, query)
tables

```
```{r}
library(ggplot2)

ggplot(data = tables, aes(x = Month, y = TotalQuantity, fill = ProductName)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Month", y = "Total Quantity", title = "Total Quantity of Products Sold per Month for each Product Name") +
  theme(plot.title = element_text(hjust = 0.5))
```
7. Consult the provider that supplies the most products. Discuss your results with the information from the supplier.
```{r}
query <- "SELECT Supplier.CompanyName, COUNT(Product.ID) AS TotalProducts
FROM Supplier
INNER JOIN Product ON Supplier.ID = Product.SupplierID
GROUP BY Supplier.ID, Supplier.CompanyName
ORDER BY TotalProducts DESC
LIMIT 10;"

tables <- dbGetQuery(con, query)
tables
```
8. Obtain a table showing the name of the product categories and the number of products that each category has. Sort by the number of products and show the first 5 items.
```{r}
query <- "SELECT Category.CategoryName, COUNT(*) as ProductCount
FROM Category
INNER JOIN Product ON Category.ID = Product.CategoryID
GROUP BY Category.CategoryName
ORDER BY ProductCount DESC
LIMIT 5;
"

tables <- dbGetQuery(con, query)
tables
```
9.Insightful Analysis
generate a pie chart of the top 5 product categories by the number of products, with each category represented by a different color.
```{r}
library(ggplot2)

result <- dbGetQuery(con, "SELECT Category.CategoryName, COUNT(*) as Count
FROM Product
INNER JOIN Category ON Product.CategoryID = Category.ID
GROUP BY Category.CategoryName
ORDER BY Count DESC
LIMIT 5")

ggplot(result, aes(x="", y=Count, fill=CategoryName)) + 
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() +
  labs(title="Top 5 Product Categories", fill="Category Name", x=NULL, y=NULL) +
  theme(legend.position="bottom")


```
